Geometric variational problems

### 2020-11-16 NSF2020/Description/specificaim1_ver2.tex
tidied it up; removed hyperlinks + pagination

%\begin{center}
%Anticipated model simulations
%\end{center}
\subsubsection{Simulations of the HAP model to estimate elastic moduli
and energy}
%To characterize material properties of a bilayer, we compare the energies and forces generated by
%a collection of amphiphilic particles against the energies and forces of a continuum bilayer.
We propose to
%
%to examine conduct calculations similar to the example of figure~\ref{fig:flattening}
%for a wide range of particle parameters and physical parameters in HARP to 
%
%{\bf (i)} Assess whether the continuum energies \eqref{ansatz3} are the same
%as the corresponding energies in the particle system for all reasonable deformation.
%If this is the case we will 
%
{\bf (i)} determine the value of effective elastic moduli $\KB$,
$\KT$, $\KG$, and $\KTH$ and then {\bf (ii)} understand how the model parameters $\rho,$ $\gamma$, and
particle shape map onto elastic moduli. 
An accurate and robust way to measure material properties is to track the evolution of the bilayer as
%
\begin{wrapfigure}[10]{r}{0.4\textwidth}
\vspace{-0.3cm}
\centerline{\includegraphics[width=0.4\textwidth]{Figures/Flattening.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
\vspace{-0.2cm}
\caption{\label{fig:flattening} \footnotesize Example of computing
  bending modulus of a lipid bilayer from particle simulation.}
\end{wrapfigure}
it relaxes from an initially non-equilibrium
configuration~\cite{PhysRevLett.117.188102}. Figure~\ref{fig:flattening}
shows an example where the initial configuration of a lipid bilayer
patch contains only splay deformation without any other components
(saddle splay, twist and tilt). As the bent membrane flattens, we find
that both the membrane self-interaction energy $\Phi$ and the elastic
energy ${\cal W}$ decrease, and because the saddle splay, twist, and
tilt stay zero, the slope gives the bending modulus in this case.

@@

Here we start with a particle-based bilayer in a specific non-equilibrium shape that involves only one of the components of the displacement in Figure~\ref{fig:deformations}, and then evolve the particle system according to the time integration for \eqref{eq:stokes}.
Because elastic properties are independent of dissipation,
we can forgo solving for fluid velocity and set the translational
velocity and angular velocity directly proportional to force and torque. 
%bilayer by steepest gradient descent 
%with respect to the HARP functional:
%\begin{equation}
%\label{SGD}   
%\frac{\dif {\bf a}_i}{\dif t} = {\bf F}_i,    \quad  \frac{\dif {\bf R}_i}{\dif t}  = {\bf R}_i A({\bf w}_i),\quad i = 1, \dots, N.
%\end{equation}
%where that particle centers ${\bf a}_i$ and rotation matrices ${\bf R}_i$ determine the particle configuraton; %
%$A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
%We solve \eqref{SGD} numerically  using an Adams–Bashforth linear multistep method. 
%The procedure involves solving \eqref{SL} based off of the current particle configuration.
%We then use the solution $u$ to calculate the collection of forces and torques \eqref{forceandtorque}, and update the particle configuration
%through \eqref{SGD}.
%
\begin{wrapfigure}[11]{l}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{Figures/SaddleSplayDiagram.pdf}}
  \vspace{-8pt}
\caption{\label{fig:saddle_splay} \footnotesize Example of determining the
  saddle splay modulus.}
\end{wrapfigure}
%
This yields a dissipation of the total potential \eqref{eq:total_poten}, which
stabilizes the evolving bilayer shape.
Therefore, we reconstruct an evolving monolayer dividing surface $\Sigma$ and director field ${\bf n}$ by
interpolating the particle centers and orientations. Using \eqref{ansatz3}, we calculate a continuum energy $\mathcal{W}$ from the interpolated shapes.

@@

We propose to conduct calculations similar to the example of Figure~\ref{fig:flattening}
for other elastic moduli such as the effective twist modulus $\KT$.
Molecular dynamics investigations find a twist modulus about 1
\kBT~\cite{LeVeWa14}, and here, the specific non-equilibrium shape consists of
a single layer of amphiphilic particles on a hydrophobic substrate as
illustrated by Figure~\ref{fig:deformations}D. Having a nonzero twist
requires nonzero tilt because the surface gradient of the lipid
director equals the second fundamental form
%, and is therefore symmetric,
whenever tilt is zero locally. The twist deformation is a fully
three-dimensional deformation. \S\ref{subsec:specific_aim_3} addresses outstanding implementation issues like three-dimensional boundary integral equation solvers.

@@

%\subsubsection{HK elastic energy form}
\noindent{\bf{HK elastic energy form:}}
The field of membrane continuum mechanics still lacks consensus as to whether \eqref{ansatz3} is the appropriate functional for bilayer energy.
\textbf{(i)} Researchers have assumed that $\KT = 0$ to effect lateral fluidity in
membranes~\cite{Hamm2000, TerziDeserno17, C9SM02079A,
PhysRevE.102.042406}. 
A value $\KT = 0$, however, makes \eqref{ansatz3} a noncoercive functional.
%and leads to the nonphysical conclusion that there are bilayers 
%with $\nabla {\bf n} \to \infty$ but $\mathcal{W} \to 0.$ 
\textbf{(ii)} Recently,~\cite{TerziDeserno17} derived a tilt curvature
term that was neglected from the HK analysis~\cite{Hamm2000}.
Later,~\cite{C9SM02079A} and~\cite{PhysRevE.102.042406} independently
identified an inconsistency in the argument used
by~\cite{TerziDeserno17} arising
from a transversal tilt invariance assumption.
In~\cite{RyKlYaCo16}, PI RR and collaborators showed that the tilt vector leads to unphysical cusps depending on how one accounts for membrane thickness.  
\textbf{(iii)} Theoretical analysis of lipid phase transitions predict a
negative saddle-splay modulus around $-8$ \kBT~\cite{SIEGEL2004366,
SIEGEL20085200} that gives rise to a larger energy barrier for
monolayer fusion than is found by experiments~\cite{FrRoPi17, Tran7106,
TerziDeserno17}.
%
%
%Researchers point out, however, that this theoretical value requires a larger energy 
%barrier for monolayer fusion than is found by experiment \cite{FrRoPi17,Tran7106,TerziDeserno17}.
%Furthermore, the stability conditions of \eqref{ansatz3} require $\KG \geq -4\min\{\KB,\KT\}$, suggesting a smaller magnitude for $\KG$.

>>>>

%\begin{center}
%Anticipated model simulations
%\end{center}
\newpage
\subsubsection{Simulations of the HAP model to estimate elastic moduli
and energy}
%To characterize material properties of a bilayer, we compare the energies and forces generated by
%a collection of amphiphilic particles against the energies and forces of a continuum bilayer.
%
%to examine conduct calculations similar to the example of figure~\ref{fig:flattening}
%for a wide range of particle parameters and physical parameters in HARP to 
%
%{\bf (i)} Assess whether the continuum energies \eqref{ansatz3} are the same
%as the corresponding energies in the particle system for all reasonable deformation.
%If this is the case we will 
%
\begin{wrapfigure}[10]{r}{0.4\textwidth}
\vspace{-0.3cm}
\centerline{\includegraphics[width=0.4\textwidth]{Figures/Flattening.pdf}}
%  \centering{\includegraphics[width=0.80\textwidth]{Figures/Deformations.pdf}}
\vspace{-0.2cm}
\caption{\label{fig:flattening} \footnotesize Example of computing
  bending modulus of a lipid bilayer from particle simulation.}
\end{wrapfigure}
We propose to {\bf (i)} determine the value of effective elastic moduli $\KB$,
$\KT$, $\KG$, and $\KTH$ and then {\bf (ii)} understand how the model parameters $\rho,$ $\gamma$, and
particle shape map onto elastic moduli. 
An accurate and robust way to measure material properties is to track the evolution of the bilayer as
it relaxes from an initially non-equilibrium
configuration~\cite{PhysRevLett.117.188102}. Figure~\ref{fig:flattening}
shows an example where the initial configuration of a bilayer
patch containing the splay deformation without any other components
(saddle splay, twist and tilt). As the bent membrane flattens, 
both the self-interaction energy $\Phi$ and the elastic
energy ${\cal W}$ decrease, and because the saddle splay, twist, and
tilt stay zero, the slope gives the bending modulus in this case.

@@

Here we start with a particle-based bilayer in a specific non-equilibrium shape that involves only one of the components of the displacement in Figure~\ref{fig:deformations}, and then evolve the particle system according to the time integration for \eqref{eq:stokes}.
Because elastic properties are independent of dissipation,
we can forgo solving for fluid velocity and set the translational
velocity and angular velocity directly proportional to force and torque. 
%bilayer by steepest gradient descent 
%with respect to the HARP functional:
%\begin{equation}
%\label{SGD}   
%\frac{\dif {\bf a}_i}{\dif t} = {\bf F}_i,    \quad  \frac{\dif {\bf R}_i}{\dif t}  = {\bf R}_i A({\bf w}_i),\quad i = 1, \dots, N.
%\end{equation}
%where that particle centers ${\bf a}_i$ and rotation matrices ${\bf R}_i$ determine the particle configuraton; %
%$A({\bf w}_i)$ is the skew-symmetric matrix with axial vector ${\bf w}_i$.
%We solve \eqref{SGD} numerically  using an Adams–Bashforth linear multistep method. 
%The procedure involves solving \eqref{SL} based off of the current particle configuration.
%We then use the solution $u$ to calculate the collection of forces and torques \eqref{forceandtorque}, and update the particle configuration
%through \eqref{SGD}.
%
This yields a dissipation of the total potential \eqref{eq:total_poten}, which
stabilizes the evolving bilayer shape.
Therefore, we reconstruct an evolving monolayer dividing surface $\Sigma$ and director field ${\bf n}$ by
interpolating the particle centers and orientations. Using \eqref{ansatz3}, we calculate a continuum energy $\mathcal{W}$ from the interpolated shapes.

@@

%
\begin{wrapfigure}[11]{l}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{Figures/SaddleSplayDiagram.pdf}}
  \vspace{-8pt}
\caption{\label{fig:saddle_splay} \footnotesize Example of determining the
  saddle splay modulus.}
\end{wrapfigure}
We propose to conduct calculations similar to the example of Figure~\ref{fig:flattening}
for other elastic moduli such as the effective twist modulus $\KT$.
Molecular dynamics investigations find a twist modulus about 1
\kBT~\cite{LeVeWa14}, and here, the specific non-equilibrium shape consists of
a single layer of amphiphilic particles on a hydrophobic substrate as
illustrated by Figure~\ref{fig:deformations}D. Having a nonzero twist
requires nonzero tilt because the surface gradient of the lipid
director equals the second fundamental form
%, and is therefore symmetric,
whenever tilt is zero locally. The twist deformation is a fully
three-dimensional deformation. \S\ref{subsec:specific_aim_3} addresses outstanding implementation issues like three-dimensional boundary integral equation solvers.

@@

%\subsubsection{HK elastic energy form}

\noindent{\bf{HK elastic energy form:}}
The field of membrane continuum mechanics still lacks consensus as to whether \eqref{ansatz3} is the appropriate functional for bilayer energy.
\textbf{(i)} Researchers have assumed that $\KT = 0$ to effect lateral fluidity in
membranes~\cite{Hamm2000, TerziDeserno17, C9SM02079A,
PhysRevE.102.042406}. 
A value $\KT = 0$, however, makes \eqref{ansatz3} a noncoercive functional.
%and leads to the nonphysical conclusion that there are bilayers 
%with $\nabla {\bf n} \to \infty$ but $\mathcal{W} \to 0.$ 
\textbf{(ii)} Recently,~\cite{TerziDeserno17} derived a tilt curvature
term that was neglected from the HK analysis~\cite{Hamm2000}.
Later,~\cite{C9SM02079A} and~\cite{PhysRevE.102.042406} independently
identified an inconsistency in the argument used
by~\cite{TerziDeserno17} arising
from a transversal tilt invariance assumption.
In~\cite{RyKlYaCo16}, PI RR and collaborators showed that the tilt vector leads to unphysical cusps depending on how one accounts for membrane thickness.  
\textbf{(iii)} Theoretical analysis of lipid phase transitions predict a
negative saddle-splay modulus around $-8$ \kBT~\cite{SIEGEL2004366,
SIEGEL20085200} that gives rise to a larger energy barrier for
monolayer fusion than is found by experiments~\cite{FrRoPi17, Tran7106,
TerziDeserno17}.
%
%
%Researchers point out, however, that this theoretical value requires a larger energy 
%barrier for monolayer fusion than is found by experiment \cite{FrRoPi17,Tran7106,TerziDeserno17}.
%Furthermore, the stability conditions of \eqref{ansatz3} require $\KG \geq -4\min\{\KB,\KT\}$, suggesting a smaller magnitude for $\KG$.
